---
title: "Tables 2, 3, and Supple Table 2"
author: "R Goto"
date: "2023-11-27"
output: html_document
---
## 4.1. Settings
```{r eval=FALSE}
library(tidyverse)
library(summarytools)
library(survey)
library(srvyr)
```

## 4.2. Dataset
- open data of cleaned 60 food consumption per day per AFE and weighting factors
```{r eval=FALSE}
f60afec <- read_csv("f60afec.csv")  # cleaned data of 60 food consumption per day per AFE
dataA <- read_csv("dataA.csv")  # data including weighting factors
```
## 4.3. the estimated consumption per day per AFE of the three food vehicles nationally, by urban/rural and four strata (Tables 2 and 3)
### 4.3.1. maize flour
```{r eval=FALSE}
# select maize flour consumption (#105)
maizef <- f60afec %>% 
  filter(itemcode==105) 

dataAw <- dataA %>% 
  select(y4_hhid, domain, strataid, clusterid, clustertype, y4_weights)  # organise dataA with weighting factors

write_csv(dataAw, "dataAw.csv")

maizefA <- left_join(maizef, dataAw, by = 'y4_hhid')   # merge with weighting factor

# weighting
maizef_w <- maizefA %>%
  as_survey_design(ids = clusterid, 
                   weights = y4_weights, 
                   strata = strataid,
                   nest=TRUE)

options(scipen = 4, digits=4)
# in national
# get estimated proportion of HHs consumed maize flour
maizef_wpr <- maizef_w %>%
  group_by(cons_yn) %>% 
  summarise(prop=survey_prop())

# get estimated maize consumption amount in median and the interquartile range (IQR)
maizef_wmd <- maizef_w %>%
  summarise(md = survey_median(g_afec, na.rm=TRUE),
            qt = survey_quantile(g_afec, c(0.25, 0.75), na.rm=TRUE))

# by urban/rural
maizef_wprur <- maizef_w %>%
  group_by(clustertype, cons_yn) %>% 
  summarise(prop=survey_prop())

maizef_wmdur <- maizef_w %>%
  group_by(clustertype) %>% 
  summarise(md = survey_median(g_afec, na.rm=TRUE),
            qt = survey_quantile(g_afec, c(0.25, 0.75), na.rm=TRUE))

# by four strata (domain)
maizef_wprst <- maizef_w %>%
  group_by(domain, cons_yn) %>% 
  summarise(prop=survey_prop())

maizef_wmdst <- maizef_w %>%
  group_by(domain) %>% 
  summarise(md = survey_median(g_afec, na.rm=TRUE),
            qt = survey_quantile(g_afec, c(0.25, 0.75), na.rm=TRUE))

```

### 4.3.1. wheat flour
- wheat flour (#1081) and added the wheat flour containing in;
buns, cakes and biscuits (#110) - 84.5% containing wheat flour
breads (#109) - 86.0%
sweets (#302) - 82.0%
macaroni and spaghetti (#111) - 100%
```{r eval=FALSE}
wheatf <- f60afec %>% 
  filter(itemcode == 1081) %>% 
  mutate(twf=g_afec) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, twf)

wbuns <- f60afec %>% 
  filter(itemcode == 110) %>% 
  mutate(twf=g_afec*0.845) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, twf)
  
wbread <- f60afec %>% 
  filter(itemcode == 109) %>% 
  mutate(twf=g_afec*0.860) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, twf)

wsweets <- f60afec %>% 
  filter(itemcode == 302) %>% 
  mutate(twf=g_afec*0.820) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, twf)

wmacaro <- f60afec %>% 
  filter(itemcode == 111) %>% 
  mutate(twf=g_afec) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, twf)

# bind rows

dftwf <- bind_rows(wheatf, wbuns, wbread, wsweets, wmacaro)

# calculate total wheat flour consumption by HHs

twf_sum <- dftwf %>% 
  group_by(y4_hhid) %>% 
  summarise(twheat = sum(twf, na.rm=TRUE),
            twhcount = sum(cons_y, na.rm = TRUE)) %>% 
  mutate(twheat=if_else(twheat>0, twheat, NA)) %>% 
  mutate(twfcons=if_else(twhcount>0, 1, 0)) %>% 
  select(-twhcount)

wheatfA <- left_join(twf_sum, dataAw, by = 'y4_hhid')   # merge with weighting factor

# weighting
wheatf_w <- wheatfA %>%
  as_survey_design(ids = clusterid, 
                   weights = y4_weights, 
                   strata = strataid,
                   nest=TRUE)

options(scipen = 4, digits=4)
# in national
# get estimated proportion of HHs consumed total wheat flour
wheatf_wpr <- wheatf_w %>%
  group_by(twfcons) %>% 
  summarise(prop=survey_prop())

# get estimated maize consumption amount in median and the interquartile range (IQR)
wheatf_wmd <- wheatf_w %>%
  summarise(wheatf_md = survey_median(twheat, na.rm=TRUE),
            wheatf_qt = survey_quantile(twheat, c(0.25, 0.75), na.rm=TRUE))

# by urban/rural
wheatf_wprur <- wheatf_w %>%
  group_by(clustertype, twfcons) %>% 
  summarise(prop=survey_prop())

wheatf_wmdur <- wheatf_w %>%
  group_by(clustertype) %>% 
  summarise(md = survey_median(twheat, na.rm=TRUE),
            qt = survey_quantile(twheat, c(0.25, 0.75), na.rm=TRUE))

# by four strata (domain)
wheatf_wprst <- wheatf_w %>%
  group_by(domain, twfcons) %>% 
  summarise(prop=survey_prop())

wheatf_wmdst <- wheatf_w %>%
  group_by(domain) %>% 
  summarise(md = survey_median(twheat, na.rm=TRUE),
            qt = survey_quantile(twheat, c(0.25, 0.75), na.rm=TRUE))
```


### 4.3.2. cooking oils
- cooking oils (#1001) and added cooking oils containing in;
buns, cakes and biscuits (#110) - 7.1% containing oils
breads (#109) - 4.9%
sweets (#302) - 12.2%
```{r eval=FALSE}
oil <- f60afec %>% 
  filter(itemcode == 1001) %>% 
  mutate(toi = g_afec) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, toi)

obun <- f60afec %>% 
  filter(itemcode == 110) %>% 
  mutate(toi = g_afec*0.071) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, toi)

obread <- f60afec %>% 
  filter(itemcode == 109) %>% 
  mutate(toi = g_afec*0.049) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, toi)

osweets <- f60afec %>% 
  filter(itemcode == 110) %>% 
  mutate(toi = g_afec*0.122) %>% 
  mutate(cons_y = if_else(cons_yn == 1, 1, NA)) %>% 
  select(y4_hhid, cons_y, toi)

# bind rows

dftoi <- bind_rows(oil, obun, obread, osweets)

# calculate total wheat flour consumption by HHs

toi_sum <- dftoi %>% 
  group_by(y4_hhid) %>% 
  summarise(toil = sum(toi, na.rm=TRUE),
            toicount = sum(cons_y, na.rm = TRUE)) %>% 
  mutate(toil=if_else(toil>0, toil, NA)) %>% 
  mutate(toilcons=if_else(toicount>0, 1, 0)) %>% 
  select(-toicount)

oilA <- left_join(toi_sum, dataAw, by = 'y4_hhid')   # merge with weighting factor

# weighting
oil_w <- oilA %>%
  as_survey_design(ids = clusterid, 
                   weights = y4_weights, 
                   strata = strataid,
                   nest=TRUE)

options(scipen = 4, digits=4)
# in national
# get estimated proportion of HHs consumed total wheat flour
oil_wpr <- oil_w %>%
  group_by(toilcons) %>% 
  summarise(prop=survey_prop())

# get estimated maize consumption amount in median and the interquartile range (IQR)
oilf_wmd <- oil_w %>%
  summarise(md = survey_median(toil, na.rm=TRUE),
            qt = survey_quantile(toil, c(0.25, 0.75), na.rm=TRUE))

# by urban/rural
oilf_wprur <- oil_w %>%
  group_by(clustertype, toilcons) %>% 
  summarise(prop=survey_prop())

oilf_wmdur <- oil_w %>%
  group_by(clustertype) %>% 
  summarise(md = survey_median(toil, na.rm=TRUE),
            qt = survey_quantile(toil, c(0.25, 0.75), na.rm=TRUE))

# by four strata (domain)
oil_wprst <- oil_w %>%
  group_by(domain, toilcons) %>% 
  summarise(prop=survey_prop())

oil_wmdst <- oil_w %>%
  group_by(domain) %>% 
  summarise(md = survey_median(toil, na.rm=TRUE),
            qt = survey_quantile(toil, c(0.25, 0.75), na.rm=TRUE))
```


## 4.4. the estimated consumption per day per AFE by 60 food items (Summplementory Table 2)
```{r eval=FALSE}
# marge with weighting fanctions
f60A <- left_join(f60afec, dataAw, by = 'y4_hhid')   # merge with weighting factor

# weighting
f60A_w <- f60A %>%
  as_survey_design(ids = clusterid, 
                   weights = y4_weights, 
                   strata = strataid,
                   nest=TRUE)

# get estimated proportion of HHs consumed 60 food items
f60_wpr <- f60A_w %>%
  group_by(itemcode, itemname, cons_yn) %>% 
  summarise(prop=survey_prop()) %>% 
  filter(cons_yn==1) %>% 
  arrange(desc(prop)) %>% 
  mutate(per=prop*100) %>% 
  mutate(per = round(per, 1))

write_csv(f60_wpr, "f60_wpr.csv")

# get estimated food consumption amount in median and the interquartile range (IQR)
f60_wmd <- f60A_w %>%
  group_by(itemcode) %>% 
  summarise(md = survey_median(g_afec, na.rm=TRUE),
            qt = survey_quantile(g_afec, c(0.25, 0.75), na.rm=TRUE))

f60_sum <- left_join(f60_wpr, f60_wmd, by = "itemcode")

names(f60_sum)

supple2 <- f60_sum %>% 
  select(itemcode, itemname, prop, md, qt_q25, qt_q75)

write_csv(supple2, "supple2.csv")

```








