---
title: "estimation of micronutrient intakes"
author: "R Goto"
date: "2023-11-30"
output: html_document
---
## 5.1. Settings
```{r eval=FALSE}
library(tidyverse)
library(summarytools)
library(survey)
library(srvyr)
```

## 5.2. Dataset
- open data of cleaned 60 food consumption per day per AFE and weighting factors
```{r eval=FALSE}
f60afec <- read_csv("f60afec.csv")  # cleaned data of 60 food consumption per day per AFE
dataAw <- read_csv("dataAw.csv")  # data including weighting factors
TZFCT3 <- read_csv("TFNC_NCT_NTPS20_v.3.0.0.csv") # organised food composition table (FCT) for 60 food items
```

## 5.3. Calculate estimated micronutrient intakes
```{r eval=FALSE}
# organise FCT
names(TZFCT3)
TZFCT3 <- TZFCT3 %>% 
  rename("itemcode"="item_id",
         "itemname"="item_desc",
         "VAmcg"="VITA_RAEmcg",
         "FOLmcg"="FOLmcg_standardised",
         "VB12mcg"="VITB12mcg")

TZFCT3 <- TZFCT3 %>% 
  select(itemcode, FEmg, ZNmg, VAmcg, FOLmcg, VB12mcg)

write_csv(TZFCT3, 'TZFCT3.csv')

# merge TZFCT with estimated food consumption in g
nf60afec <- left_join(f60afec, TZFCT3, by = "itemcode")

# calculate estimated micronutrient intakes per day per AFE in HHs considering that FCT values are calculated nutrient amount per 100g

options(scipen = 10, digits=2)

nf60afec <- nf60afec %>% 
  mutate(feafe = g_afec*FEmg*0.01,
         znafe = g_afec*ZNmg*0.01,
         vaafe = g_afec*VAmcg*0.01,
         folafe = g_afec*FOLmcg*0.01,
         vb12afe = g_afec*VB12mcg*0.01)

# calculate sum of nutrients from 60 food items by HHs
nafet <- nf60afec %>% 
  group_by(y4_hhid) %>% 
  summarise(feafet=sum(feafe, na.rm=TRUE),
            znafet=sum(znafe, na.rm=TRUE),
            vaafet=sum(vaafe, na.rm=TRUE),
            folafet=sum(folafe, na.rm=TRUE),
            vb12afet=sum(vb12afe, na.rm=TRUE))

write_csv(nafet, "nafet.csv")

# weighting
nmA <- left_join(nafet, dataAw, by = 'y4_hhid')   # merge with weighting factor

nmA_w <- nmA %>%
  as_survey_design(ids = clusterid, 
                   weights = y4_weights, 
                   strata = strataid,
                   nest=TRUE)

# get estimated micronutrient intakes by 60 food items in median and the interquartile range (IQR)
# in national (Table 4)
options(scipen = 4, digits=4)
nmA_wmd <- nmA_w %>%
  summarise(femd = survey_median(feafet, na.rm=TRUE),
            feqt = survey_quantile(feafet, c(0.25, 0.75), na.rm=TRUE),            ,
            znmd = survey_median(znafet, na.rm=TRUE),
            znqt = survey_quantile(znafet, c(0.25, 0.75), na.rm=TRUE),
            vamd = survey_median(vaafet, na.rm=TRUE),
            vaqt = survey_quantile(vaafet, c(0.25, 0.75), na.rm=TRUE),
            folmd = survey_median(folafet, na.rm=TRUE),
            folqt = survey_quantile(folafet, c(0.25, 0.75), na.rm=TRUE),
            vb12md = survey_median(vb12afet, na.rm=TRUE),
            vb12qt = survey_quantile(vb12afet, c(0.25, 0.75), na.rm=TRUE)
            )

# by urban/rural
nmA_wmdur <- nmA_w %>%
  group_by(clustertype) %>% 
  summarise(femd = survey_median(feafet, na.rm=TRUE),
            feqt = survey_quantile(feafet, c(0.25, 0.75), na.rm=TRUE),
            znmd = survey_median(znafet, na.rm=TRUE),
            znqt = survey_quantile(znafet, c(0.25, 0.75), na.rm=TRUE),
            vamd = survey_median(vaafet, na.rm=TRUE),
            vaqt = survey_quantile(vaafet, c(0.25, 0.75), na.rm=TRUE),
            folmd = survey_median(folafet, na.rm=TRUE),
            folqt = survey_quantile(folafet, c(0.25, 0.75), na.rm=TRUE),
            vb12md = survey_median(vb12afet, na.rm=TRUE),
            vb12qt = survey_quantile(vb12afet, c(0.25, 0.75), na.rm=TRUE)
            )

# by strata
nmA_wmddo <- nmA_w %>%
  group_by(domain) %>% 
  summarise(femd = survey_median(feafet, na.rm=TRUE),
            feqt = survey_quantile(feafet, c(0.25, 0.75), na.rm=TRUE),
            znmd = survey_median(znafet, na.rm=TRUE),
            znqt = survey_quantile(znafet, c(0.25, 0.75), na.rm=TRUE),
            vamd = survey_median(vaafet, na.rm=TRUE),
            vaqt = survey_quantile(vaafet, c(0.25, 0.75), na.rm=TRUE),
            folmd = survey_median(folafet, na.rm=TRUE),
            folqt = survey_quantile(folafet, c(0.25, 0.75), na.rm=TRUE),
            vb12md = survey_median(vb12afet, na.rm=TRUE),
            vb12qt = survey_quantile(vb12afet, c(0.25, 0.75), na.rm=TRUE)
            )

```

## 5.4. Calculate nutrient gap (Table 4)

### 5.4.1. calculate the gap between estimated nutrient intakes and 25 percentile of the population
# the 25 percentile of each nutrient was calculated in 5.3 and manually calculate the differences between the values and H-AR
# H-AR for females aged 18-50 yrs (Allen et al., 2020): iron 22.4mg, zinc 10.2mg, VA 490mcg, folate 250mcg, VB12 2.0mcg
- for example, 25 percentile of the iron intake nationally was 14.9mg/d/AFE. H-AR for female 18-50 yrs of age was 22.4mg, therefore the gap is 14.9-22.4=-7.5mg

### 5.5. calculate the prevalance of micronutrient inadequecy in no fortification scenario (Table 5)
```{r eval=FALSE}
nafet <- nafet %>% 
  mutate(znadq=if_else(znafet<10.2, "inadequate", "adequate"),
         vaadq=if_else(vaafet<490, "inadequate", "adequate"),
         foladq=if_else(folafet<250, "inadequate", "adequate"),
         vb12adq=if_else(vb12afet<2.0, "inadequate", "adequate"))

# weighting 
nafet_w <- left_join(nafet, dataAw, by = 'y4_hhid')   # merge with weighting factor

nafet_w <- nafet_w %>%
  as_survey_design(ids = clusterid, 
                   weights = y4_weights, 
                   strata = strataid,
                   nest=TRUE)

# proportion at the zinc inadequacy
options(digits=3)
# nationally
znadqn <- nafet_w %>%
  group_by(znadq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by urban/rural
znadqur <- nafet_w %>%
  group_by(clustertype, znadq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by strata
znadqdo <- nafet_w %>%
  group_by(domain, znadq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))

# proportion at the VA inadequacy
# nationally
vaadqn <- nafet_w %>%
  group_by(vaadq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by urban/rural
vaadqur <- nafet_w %>%
  group_by(clustertype, vaadq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by strata
vaadqdo <- nafet_w %>%
  group_by(domain, vaadq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))

# proportion at the folate inadequacy
# nationally
foladqn <- nafet_w %>%
  group_by(foladq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by urban/rural
foladqur <- nafet_w %>%
  group_by(clustertype, foladq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by strata
foladqdo <- nafet_w %>%
  group_by(domain, foladq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))

# proportion at the VB12 inadequacy
# nationally
vb12adqn <- nafet_w %>%
  group_by(vb12adq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by urban/rural
vb12adqur <- nafet_w %>%
  group_by(clustertype, vb12adq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))
# by strata
vb12adqdo <- nafet_w %>%
  group_by(domain, vb12adq) %>% 
  summarise(prop=survey_prop( , vartype = c("ci"),  level = 0.95))

```
















